### 2022-08-23 main/chapter_searches.tex
some text on use of bayes rule

We finally have the four parts of `Bayesian' data analysis:
the two inputs \textbf{likelihood} and \textbf{prior},
and two outputs \textbf{posterior} and \textbf{evidence}.
These relate through Bayes' rule, which is a consequence of the
produce rule of Equation~\ref{eqn:searches_product_rule} and the
commutativity of `and' --- that `$a$ and $b$' is equivalent to
`$b$ and $a$';
the joint probability of data and parameters has two identical
forms, $\prob{d, x}{h}$ and $\prob{x, d}{h}$, that each factor into two
pieces:
\begin{align}
\prob{d}{x, h}\times \prob{x}{h} &= \prob{x}{d, h}\times \prob{d}{h},
\intertext{all of which we have named:}
\underbrace{L(x)}_\textbf{likelihood}
\times
~\,
\overbrace{\pi(x)}^\textbf{prior}
~\,
&=
\overbrace{p(x)}^\textbf{posterior}
\times
\underbrace{Z}_\textbf{evidence}
.
% \textbf{likelihood} \times \textbf{prior}
% &= \textbf{posterior} \times \textbf{evidence} .\nonumber
\end{align}
This is (the infamous) Bayes' rule, derived as a consequence of the
symmetry of logical `and' and the product rule of probability
algebra.
All influence from data is conducted through the likelihood function
\TODO{here state how we should use this}.

The infamy of Bayes' rule stems from a fixation on the posterior distributions
from unjustified prior assignments, which has been present since its earliest
classical developments by Bayes and
Laplace~\cite{bayes1763lii, laplace1774stigler}.
Both Bayes and Laplace assume prior densities that are uniform in their chosen
coordinates, a practice that is fairly criticized as arbitrary
\TODO{zero evidence because for uniform on the real line --- a normalized prior
with any inkling of the scale of the actual parameter will be infinitely
preferred in its evidence.}

% likelihood ratio
\begin{equation}
\hphantom{\quad \quad \mid\mid \mathrm{context}}
\frac{\prob{S}{d}}{\prob{B}{d}}
=
\!\!
\underbrace{\frac{\prob{d}{S}}{\prob{d}{B}}}_\textrm{likelihood ratio}
\!\!\times~
\frac{\probc{S}}{\probc{B}}
\quad \quad \mid\mid \mathrm{context}
\end{equation}

Probabilistic learning proceeds as follows:
\begin{enumerate}
\item Predict new data $d$ with probability $\prob{d}{\ldots, h}$.
\item Update your predictions with the observed data to $\prob{x}{d, \ldots, h}$.
\item Go to 1.
\end{enumerate}

@@

% summary

Figure~\ref{fig:searches_baton_roue_bayes}

>>>>

We finally have the four parts of `Bayesian' data analysis:
the two inputs \textbf{likelihood} and \textbf{prior},
and two outputs \textbf{posterior} and \textbf{evidence}.
These relate through Bayes' rule, which is a consequence of the
produce rule of Equation~\ref{eqn:searches_product_rule} and the
commutativity of `and' --- that `$a$ and $b$' is equivalent to
`$b$ and $a$';
the joint probability of data and parameters has two identical
forms, $\prob{d, x}{h}$ and $\prob{x, d}{h}$, that each factor into two
pieces:
\begin{align}
\prob{d}{x, h}\times \prob{x}{h} &= \prob{x}{d, h}\times \prob{d}{h},
\intertext{all of which we have named:}
\underbrace{L(x)}_\textbf{likelihood}
\times
~\,
\overbrace{\pi(x)}^\textbf{prior}
~\,
&=
\overbrace{p(x)}^\textbf{posterior}
\times
\underbrace{Z}_\textbf{evidence}
.
% \textbf{likelihood} \times \textbf{prior}
% &= \textbf{posterior} \times \textbf{evidence} .\nonumber
\end{align}
This is (the infamous) Bayes' rule, derived as a consequence of the
symmetry of logical `and' and the product rule of probability
algebra.

How Bayes' rule \emph{should} be used in scientific data analysis is to extract
useful information from the data.
If the likelihood function itself is interpretable, then it should be reported
directly.
When it is not, perhaps due to a complicated parametric modelling, then one
should abstract those parameters away by reporting evidence values for
predictive models, which must of course be clearly stated and interpretable if
they are to be useful.
Data act only through the likelihood function, and reporting a posterior
can act to obfuscate those data.

Posterior probabilities are, however, useful when digesting data to make
inferences about the world.
Reporting likelihoods and evidence values allows readers to assign priors as
they wish.
For model comparison between background $h=B$ and signal $h=S$, this is clearly
expressed in the ratio form
% likelihood ratio
\begin{equation}
\hphantom{\quad \quad \mid\mid \mathrm{context}}
\frac{\prob{S}{d}}{\prob{B}{d}}
=
\!\!
\frac{\prob{d}{S}}{\prob{d}{B}}
\!\!\times~
\frac{\probc{S}}{\probc{B}}
\quad \quad \mid\mid \mathrm{context}
,
\end{equation}
where we use the double bar, $\mid\mid$, to refer to any contextual information
a reader might wish to include.
Whenever the evidence ratio is independent of that context, it factors out from
the prior (odds) ratio.
In this sense, likelihood (or evidence) ratios are scale factors by which prior
odds are updated.

The infamy of Bayes' rule stems from a fixation on the posterior distributions
from unjustified prior assignments.
Classical works by Bayes and Laplace~\cite{bayes1763lii, laplace1774stigler}
began this focus;
both assume prior densities that are uniform in their chosen coordinates, a
practice that can yield practical results in posterior inferences.
But for our purposes of reporting scientific data, such practices are fairly
criticized as arbitrary when they influce our results.
\TODO{references for definition of Bayesian analysis as computation of
posterior without even giving a symbool to the evidence}
Figure~\ref{fig:searches_baton_roue_bayes}

\TODO{zero evidence because for uniform on the real line --- a normalized prior
with any inkling of the scale of the actual parameter will be infinitely
preferred in its evidence.}
oceeds as follows:
\begin{enumerate}
\item Predict new data $d$ with probability $\prob{d}{\ldots, h}$.
\item Update your predictions with the observed data to $\prob{x}{d, \ldots, h}$.
\item Go to 1.
\end{enumerate}

@@
